Article data analysis 2

### 2019-03-15 notebooks/article_notebook2.Rmd
removed clustering analysis

```{r load libraries, message=FALSE, warning=FALSE}
# Libraries
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr, ggplot2, tidyr, cowplot, viridis, forcats, radiant.data, purrr, phangorn, gridExtra, ggtree, htmlTable, ape, cluster, tibble, Biostrings, seqinr, stringr)
```

```{r Import data, include = FALSE}
# Collapses data frame to unique lines while ignoring NA's
func_paste <- function(x) paste(unique(x[!is.na(x)]), collapse = ", ")

@@

isolate_data <- read.table("../data/isolate_data.txt",
                           sep = "\t",
                           header = T,
                           stringsAsFactors = F) %>%
  filter(id %not_in% c("2016-17-565-1-S",
                       "2016-02-428-2-S",
                       "2016-02-486-2-S",
                       "2016-02-732-2-S",
                       "2014-01-1741-1-S"))

@@

no_of_reads <- as.data.frame(rbind(as.matrix(two_reads), as.matrix(four_reads))) %>%
  left_join(id_data, by = "ref") %>%
  select(id, everything(), -ref) %>%
  filter(id %not_in% c("2016-17-565-1-S",
                       "2016-02-428-2-S",
                       "2016-02-486-2-S",
                       "2016-02-732-2-S",
                       "2014-01-1741-1-S"))

mic_data <- read.table("../data/mic_data.txt",
                           sep = "\t",
                           header = T,
                           stringsAsFactors = F) %>%
  filter(id %not_in% c("2016-17-565-1-S",
                       "2016-02-428-2-S",
                       "2016-02-486-2-S",
                       "2016-02-732-2-S",
                       "2014-01-1741-1-S"))

@@

```{r ARIBA analysis, include = FALSE}
megares_report_loc <- "../data/ariba/megares_results"
resfinder_report_loc <- "../data/ariba/resfinder_results"
ac_genes <- c("qnr","oqx","qep","mfp","mcb")
mut_genes <- c("gyr","par","marR","marA","acrB","soxR","acrR","nor","ydh","sdi","rpoB")

# adjust parameters for filtering
if (length(ac_genes) == 1) {
  if (grepl("all", ac_genes, ignore.case = TRUE) == TRUE) {
    ac_genes <- "ALL"
    } else {
      ac_genes <- ac_genes
    }
}

if (length(mut_genes) == 1) {  
  if (grepl("all", mut_genes, ignore.case = TRUE) == TRUE) {
    mut_genes <- "ALL"
    } else {
      mut_genes <- mut_genes
    }
}
# -------------------------------------- Libraries

if (!require("pacman")) install.packages("pacman")
pacman::p_load(ggplot2,dplyr,tidyr,gridExtra,grid,
               forcats,purrr,stringr,kableExtra,
               knitr,IRdisplay,reprex,svglite)

# -------------------------------------- Functions

func_paste2 <- function(x) paste(unique(x[!is.na(x)]), collapse = ",")

# Confidence interval function
get_binCI <- function(x, n) as.numeric(setNames(binom.test(x,n)$conf.int*100,
                                                c("lwr", "upr")))

@@

# Import ariba data from report.tsv from chosen database used in ariba
get_ariba_data <- function(filepath) {
  files <- file_names(filepath)

  data_list <- lapply(files,
                      FUN = function(file) {
                        read.delim(
                          paste0(filepath, "/", file),
                          stringsAsFactors = F,
                          header = TRUE,
                          sep = "\t"
                        )
                      })

  names(data_list) <- files
  data <- bind_rows(lapply(data_list, function(x) map(x, as.character)), .id = "ref")
  return(data)
}

# Corrects the gene names found in the "cluster" column
fix_gene_names <- function(df) {
  genes <- unique(df$ref_name)
  new_names <- gsub("^(.*?)\\..*", "\\1", genes)
  new_names <- gsub("_", "", new_names, fixed = T)
  new_names <- gsub("-", "", new_names, fixed = T)

  gene_names <- c()
  for (i in new_names) {
    p <- paste(tolower(substring(i, 1,3)), substring(i, 4), sep = "", collapse = " ")
    gene_names <- c(gene_names,p)
  }
  df2 <- data.frame(genes,gene_names) %>%
    mutate(genes = as.character(genes)) %>%
    dplyr::rename(ref_name = genes)

  df <- df %>%
    left_join(df2, by = "ref_name") %>%
    mutate(gene_names = as.character(gene_names),
           ref = gsub("(.*?)_amr_report.tsv", "\\1", ref))

@@

# Function that handles megares data and returns a data frame with information on whether a gene
# is mutated or not. Includes control for QRDR in gyrA.
create_mut_table <- function(df) {
  df <- df %>%
    mutate(ref_ctg_change = ifelse(ref_ctg_change == ".", "", ref_ctg_change),
           ref_nt = ifelse(ref_nt == ".", "", ref_nt),
           ctg_nt = ifelse(ctg_nt == ".", "", ctg_nt),
           ref_ctg_nt_change = paste(ref_nt, ctg_nt, sep = "-"),
           mutation = paste(ref_ctg_change, ref_ctg_nt_change, sep = "/"),
           mutation = ifelse(mutation == "/-", "0", mutation)) %>%
    select(ref, gene_names, flag, mutation) %>%
    filter(flag %in% flag_selection) %>%
    mutate(id = 1:n()) %>%
    spread(gene_names, mutation) %>%
    group_by(ref) %>%
    summarise_all(funs(func_paste)) %>%
    select(-c(id, flag)) %>%
    gather(gene, mut, -ref) %>%
    mutate(mut = ifelse(mut == "" | mut == "." | is.na(mut) == TRUE, 0, mut),
           result = ifelse(mut != 0, 1, 0),
           result = as.integer(result),
           type = "mut",
           nt_change = gsub("[^,]+?/([^,]+?)", "\\1", mut),
           mut = gsub("([^/]+)/[^,]+", "\\1", mut))
  return(df)
}

# Corrects or mutations in QRDR for gyrA, gyrB, parC and parE genes.
# The function returns columns of 1/0 values for whether or not the 
# mutations reported by ARIBA is within the QRDR in the gene:
# gyrA: AA 67 - 106
# gyrB: AA 333 - 481
# ParC: AA 51 - 170
# parE: AA 366 - 523
# Source: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC1142146/pdf/cjvr68pg229.pdf
fix_gyr_par_results <- function(df) {
  df <- df %>%
    mutate(gyrA_result = mut %>% # control for mutation within QRDR for gyrA
             str_extract_all("\\d+") %>% # from reprex package
             map(as.integer) %>% # converts all to integer
             map_lgl(~ any(.x >= 67L & .x <= 106L)), # returns TRUE/FALSE whether value is within range or not
           gyrA_result = if_else(gene != "gyrA", NA, gyrA_result), # filters out results for all other genes
           gyrA_result = as.integer(gyrA_result),
           gyrB_result = mut %>% # control for mutation within QRDR for gyrB
             str_extract_all("\\d+") %>%
             map(as.integer) %>%
             map_lgl(~ any(.x >= 333L & .x <= 481L)),
           gyrB_result = if_else(gene != "gyrB", NA, gyrB_result),
           gyrB_result = as.integer(gyrB_result),
           parC_result = mut %>% # control for mutation within QRDR for parC
             str_extract_all("\\d+") %>% 
             map(as.integer) %>% 
             map_lgl(~ any(.x >= 51L & .x <= 170L)),
           parC_result = if_else(gene != "parC", NA, parC_result),
           parC_result = as.integer(parC_result),
           parE_result = mut %>% # control for mutation within QRDR for parE
             str_extract_all("\\d+") %>% 
             map(as.integer) %>% 
             map_lgl(~ any(.x >= 366L & .x <= 523L)),
           parE_result = if_else(gene != "parE", NA, parE_result),
           parE_result = as.integer(parE_result)) %>%
    mutate(result_gyr_par = case_when(gene == "gyrA" ~ gyrA_result,
                                      gene == "gyrB" ~ gyrB_result,
                                      gene == "parC" ~ parC_result,
                                      gene == "parE" ~ parE_result)) %>%
    mutate(result_total = ifelse(gene %in% c("gyrA","gyrB","parC","parE"), result_gyr_par, result)) %>%
    select(-c(gyrA_result,
              gyrB_result,
              parC_result,
              parE_result,
              result_gyr_par,
              result))
  return(df)
}

# Function that handles resfinder data and returns a data frame with 
# presence/absence for acquired genes                           
create_acquired_table <- function(df) {
  df <- df %>%
    select(ref, gene_names, flag, ref_ctg_change) %>%
    mutate(id = 1:n()) %>%
    filter(flag %in% flag_selection) %>%
    spread(gene_names, ref_ctg_change) %>%
    group_by(ref) %>%
    summarise_all(funs(func_paste)) %>%
    mutate(ref = gsub("^(.*?)_amr_report.tsv$", "\\1", ref)) %>%
    select(-c(id, flag)) %>%
    gather(gene, result,-ref) %>%
    mutate(result_total = ifelse(result == "", 0, 1),
           result_total = as.character(result_total),
           type = "gene") %>%
    select(-result)
  return(df)
}

# Function that returns a filtered dataframe based on the string "genes"
filter_mut_table <- function(df) {
  report_genes <- unique(df$gene)
  grep_genes <- c()
  for (gene in mut_genes) {
    for (g in report_genes) {
      if (grepl(gene, g, ignore.case = T) == TRUE) {
        grep_genes <- c(grep_genes, g)
      }
    }
  }
  df <- df %>%
    filter(gene %in% grep_genes) %>%
    mutate(gene = as.character(gene))
  return(df)
}

# Function that returns a filtered dataframe based on the string "acquired_genes"
filter_acquired_table <- function(df) {
  report_genes <- unique(df$gene)
  grep_genes <- c()
  for (gene in ac_genes) {
    for (g in report_genes) {
      if (grepl(gene, g, ignore.case = T) == TRUE) {
        grep_genes <- c(grep_genes, g)
      }
    }
  }
  df <- df %>%
    filter(gene %in% grep_genes) %>%
    mutate(gene = gsub("_", "", gene),
           gene = as.character(gene))
  return(df)
}

# calculates percentage of present/absent mutations and genes
calc_stats <- function(df) {
  df <- df %>%
    group_by(gene, result_total) %>%
    dplyr::count() %>%
    ungroup() %>%
    mutate(result_total = if_else(result_total == 1, "Present", "Absent")) %>%
    spread(result_total, n, fill = 0) %>%
    rowwise() %>%
    mutate(Total = Present + Absent,
           Percent = round(Present/Total*100, 1))
  return(df)
}

# calculates how many mutations are present in the genes
# gyrA, gyrB, parC and parE
calc_no_of_mut <- function(df) {
  df1 <- df %>%
    filter(gene %in% c("gyrA","parC","gyrB","parE")) %>%
    mutate(entries = sapply(strsplit(.$mut, ","), FUN = function(x) {length(x)})) %>%
    separate(mut, into = as.character(c(1:max(.$entries)))) %>%
    select(-entries) %>%
    gather(id, mut, -c(ref, gene, result, type))

  df2 <- fix_gyr_par_results(df1)

  df3 <- df2 %>%
    mutate(test = ifelse(mut != "0" & result_total == 0, 0, 1)) %>%
    filter(test == 1) %>%
    spread(gene, mut) %>%
    group_by(ref) %>%
    summarise_all(funs(func_paste)) %>%
    mutate_at(.vars = vars(c("gyrA","gyrB","parC","parE")),
              .funs = function(x) ifelse(x == "0", "", x)) %>%
    mutate(mut_gyrA = sapply(strsplit(.$gyrA, ","), FUN = function(x) {length(x)}),
           mut_gyrB = sapply(strsplit(.$gyrB, ","), FUN = function(x) {length(x)}),
           mut_parC = sapply(strsplit(.$parC, ","), FUN = function(x) {length(x)}),
           mut_parE = sapply(strsplit(.$parE, ","), FUN = function(x) {length(x)})) %>%
    select(-c(type, id, result_total, test)) %>%
    mutate_at(.vars = vars(c("gyrA","gyrB","parC","parE")),
              .funs = function(x) ifelse(x == "", "0", x))
  return(df3)
}

@@

# -------------------------------------- Analysis
# Import data
mut_data <- get_ariba_data(megares_report_loc)
acquired_data <- get_ariba_data(resfinder_report_loc)

@@

# Check flags
mut_flags <- check_flags(clean_mut_data) %>%
  left_join(id_data, by = "ref") %>%
  select(id, everything(), -ref) %>%
  filter(id %not_in% c("2016-17-565-1-S",
                       "2016-02-428-2-S",
                       "2016-02-486-2-S",
                       "2016-02-732-2-S",
                       "2014-01-1741-1-S"))

acquired_flags <- check_flags(clean_acquired_data) %>%
  left_join(id_data, by = "ref") %>%
  select(id, everything(), -ref) %>%
  filter(id %not_in% c("2016-17-565-1-S",
                       "2016-02-428-2-S",
                       "2016-02-486-2-S",
                       "2016-02-732-2-S",
                       "2014-01-1741-1-S"))

# Wrangle
mut_table <- create_mut_table(clean_mut_data)
mut_table_fixed <- fix_gyr_par_results(mut_table) %>%
  left_join(id_data, by = "ref") %>%
  select(id, everything(), -ref) %>%
  filter(id %not_in% c("2016-17-565-1-S",
                       "2016-02-428-2-S",
                       "2016-02-486-2-S",
                       "2016-02-732-2-S",
                       "2014-01-1741-1-S"))

mut_filtered <- filter_mut_table(mut_table_fixed)
acquired_table <- create_acquired_table(clean_acquired_data) %>%
  left_join(id_data, by = "ref") %>%
  select(id, everything(), -ref) %>%
  filter(id %not_in% c("2016-17-565-1-S",
                       "2016-02-428-2-S",
                       "2016-02-486-2-S",
                       "2016-02-732-2-S",
                       "2014-01-1741-1-S"))

@@

# Stats
if (length(mut_genes) == 1) {
  if (mut_genes == "ALL") {
    mut_stats <- calc_stats(mut_table_fixed)
    mut_report <- create_mut_report(mut_table_fixed)
    }
} else {
  mut_stats <- calc_stats(mut_filtered)
  mut_report <- create_mut_report(mut_filtered)
}

if (length(ac_genes) == 1) {
  if (ac_genes == "ALL") {
    acquired_stats <- calc_stats(acquired_table)
    acquired_report <- create_acquired_report(acquired_table)
  }
} else {
  acquired_stats <- calc_stats(acquired_filtered)
  acquired_report <- create_acquired_report(acquired_filtered)
}

mut_quant <- suppressWarnings(calc_no_of_mut(mut_table)) %>%
  left_join(id_data, by = "ref") %>%
  select(id, everything(), -ref) %>%
  filter(id %not_in% c("2016-17-565-1-S",
                       "2016-02-428-2-S",
                       "2016-02-486-2-S",
                       "2016-02-732-2-S",
                       "2014-01-1741-1-S"))

# ---------------------------------- Tables

@@

# Double-check ARIBA results with annotated sequences from Prokka
```{r, warning=FALSE, message=FALSE}
# Functions
mismatches <- function(query, ref) {
  pairwiseAlignment(ref, query, substitutionMatrix = "BLOSUM50",
                    gapOpening = 3, gapExtension = 1) %>%
    mismatchTable() %>%
    mutate(ID=names(query),
           Pos=PatternStart,
           Reference_AA=as.character(PatternSubstring),
           Sample_AA=as.character(SubjectSubstring)) %>%
    select(ID, Reference_AA, Sample_AA, Pos) %>%
    filter(Sample_AA != "X")
}

# Run mutation identification analysis on Prokka sequences and
# match to ARIBA results
gyrA_results <- bind_rows(lapply(seq_along(gyrA[-1]),
                            function(i) mismatches(gyrA[i + 1],
                                                   gyrA[1]))) %>%
  mutate(mutation = paste0(Reference_AA, Pos, Sample_AA)) %>%
  select(ID, mutation) %>%
  mutate(ID = sub(".faa", "", ID)) %>%
  dplyr::rename("ref" = ID,
                "gyrA_prokka" = mutation) %>%
  left_join(id_data, by = "ref") %>%
  select(id, gyrA_prokka) %>%
  mutate(gyrA_result = gyrA_prokka %>% # control for mutation within QRDR for gyrA
             str_extract_all("\\d+") %>% # from reprex package
             map(as.integer) %>% # converts all to integer
             map_lgl(~ any(.x >= 67L & .x <= 106L))) %>%
  filter(gyrA_result == TRUE) %>%
  group_by(id) %>%
  mutate(test = 1:n()) %>%
  ungroup() %>%
  left_join(mut_gyrA, by = c("id","test")) %>%
  select(-test) %>%
  dplyr::rename("gyrA_ariba" = value) %>%
  filter(id %not_in% c("2016-17-565-1-S",
                       "2016-02-428-2-S",
                       "2016-02-486-2-S",
                       "2016-02-732-2-S",
                       "2014-01-1741-1-S")) %>%
  select(-gyrA_result)

# Run mutation identification analysis on Prokka sequences and
# match to ARIBA results
parC_results <- bind_rows(lapply(seq_along(parC[-1]),
                                 function(i) mismatches(parC[i + 1],
                                                        parC[1]))) %>%
  mutate(Pos = Pos + 22) %>%
  mutate(mutation = paste0(Reference_AA, Pos, Sample_AA)) %>%
  select(ID, mutation) %>%
  mutate(ID = sub(".faa", "", ID)) %>%
  dplyr::rename("ref" = ID,
                "parC_prokka" = mutation) %>%
  left_join(id_data, by = "ref") %>%
  select(id, parC_prokka) %>%
  mutate(parC_result = parC_prokka %>%
             str_extract_all("\\d+") %>% 
             map(as.integer) %>% 
             map_lgl(~ any(.x >= 51L & .x <= 170L))) %>%
  filter(parC_result == TRUE) %>%
  group_by(id) %>%
  mutate(test = 1:n()) %>%
  ungroup() %>%
  left_join(mut_parC, by = c("id","test")) %>%
  select(-test) %>%
  dplyr::rename("parC_ariba" = value) %>%
  filter(id %not_in% c("2016-17-565-1-S",
                       "2016-02-428-2-S",
                       "2016-02-486-2-S",
                       "2016-02-732-2-S",
                       "2014-01-1741-1-S")) %>%
  select(-parC_result)

# Run mutation identification analysis on Prokka sequences and
# match to ARIBA results
parE_results <- bind_rows(lapply(seq_along(parE[-1]),
                                 function(i) mismatches(parE[i + 1],
                                                        parE[1]))) %>%
  #mutate(Pos = Pos + 22) %>%
  mutate(mutation = paste0(Reference_AA, Pos, Sample_AA)) %>%
  select(ID, mutation) %>%
  mutate(ID = sub(".faa", "", ID)) %>%
  dplyr::rename("ref" = ID,
                "parE_prokka" = mutation) %>%
  left_join(id_data, by = "ref") %>%
  select(id, parE_prokka) %>%
  mutate(parE_result = parE_prokka %>%
           str_extract_all("\\d+") %>% 
           map(as.integer) %>% 
           map_lgl(~ any(.x >= 366L & .x <= 523L))) %>%
  filter(parE_result == TRUE) %>%
  group_by(id) %>%
  mutate(test = 1:n()) %>%
  ungroup() %>%
  left_join(mut_parE, by = c("id","test")) %>%
  select(-test) %>%
  dplyr::rename("parE_ariba" = value) %>%
  filter(id %not_in% c("2016-17-565-1-S",
                       "2016-02-428-2-S",
                       "2016-02-486-2-S",
                       "2016-02-732-2-S",
                       "2014-01-1741-1-S")) %>%
  select(-parE_result)

# Tables
## Isolate data
```{r}
isolate_data_complete <- isolate_data %>%
  left_join(no_of_reads, by = "id")

@@

## List of genes of interest
This table lists all the genes of interest in this study, some information about them, and their accession number(s).
```{r}

acquired_ref_names <- acquired_data$ref_name
acquired_ref <- c()

for (i in ac_genes) {
  for (j in acquired_ref_names) {
    if (grepl(i,j, ignore.case = T) == TRUE) {
      acquired_ref <- c(acquired_ref, j)
    }
  }
}

acquired_ref <- unique(acquired_ref)

>>>>

```{r load libraries, message=FALSE, warning=FALSE}
# Libraries
# WARNING: This will install packages on your system if not present in library path!
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr, ggplot2, tidyr, cowplot, viridis, forcats, radiant.data, purrr, phangorn, gridExtra, ggtree, ape, cluster, tibble, Biostrings, seqinr, stringr, scales, reprex)
```

```{r Import data, include = FALSE}
# Vector of excluded samples
ex_samples <- c(
  "2016-17-565-1-S",
  "2016-02-428-2-S",
  "2016-02-486-2-S",
  "2016-02-732-2-S",
  "2014-01-1741-1-S"
  )

# Collapses data frame to unique lines while ignoring NA's
func_paste <- function(x) paste(unique(x[!is.na(x)]), collapse = ", ")

@@

isolate_data <- read.table("../data/isolate_data.txt",
                           sep = "\t",
                           header = T,
                           stringsAsFactors = F) %>%
  filter(id %not_in% ex_samples)

@@

no_of_reads <- as.data.frame(rbind(as.matrix(two_reads),
                                   as.matrix(four_reads))) %>%
  left_join(id_data, by = "ref") %>%
  select(id, everything(), -ref) %>%
  filter(id %not_in% ex_samples)

mic_data <- read.table("../data/mic_data.txt",
                           sep = "\t",
                           header = T,
                           stringsAsFactors = F) %>%
  filter(id %not_in% ex_samples)

@@

# ARIBA reports analyses
This script filters results from ARIBA reports for genes of interest
```{r ARIBA analysis}
# Report directories
megares_report_loc <- "../data/ariba/megares_results"
resfinder_report_loc <- "../data/ariba/resfinder_results"

# Genes of interest. These genes are partially matched to 
# the generated gene names from the "fix_gene_names" function
ac_genes <- c("qnr","oqx","qep","aac6")
mut_genes <- c("gyr","par","marR","marA",
               "sox","rpoB","rob","gad")

# -------------------------- Functions ----------------------------

# Paste unique cell values together, separate with ","
func_paste2 <- function(x) paste(unique(x[!is.na(x)]),
                                 collapse = ",")

# Confidence interval function
get_binCI <- function(x, n) as.numeric(
  setNames(
    binom.test(x,n)$conf.int*100, 
    c("lwr", "upr")
    )
  )

@@

# Import ariba data from report.tsv identified with file_names
get_ariba_data <- function(filepath) {
  # Identify file names in filepath
  files <- file_names(filepath)

  # Import data
  data_list <- lapply(files,
                      FUN = function(file) {
                        read.delim(
                          paste0(filepath, "/", file),
                          stringsAsFactors = F,
                          header = TRUE,
                          sep = "\t"
                        )
                      })

  # Set file names in list
  names(data_list) <- files

  # Bind to data frame and convert columns to character
  data <- bind_rows(
    lapply(
      data_list,
      function(x) map(x, as.character)
      ),
    .id = "ref")

  return(data)
}

# Corrects gene names 
fix_gene_names <- function(df) {
  # Identify unique entries in ref_name column
  genes <- unique(df$ref_name)

  # Strip away excess characters
  new_names <- gsub("^(.*?)\\..*", "\\1", genes)
  new_names <- gsub("_", "", new_names, fixed = T)
  new_names <- gsub("-", "", new_names, fixed = T)

  # make all genes lower case, with last letter 
  # upper case if four characters long
  gene_names <- c()

  for (i in new_names) {
    p <- paste(
      tolower(
        substring(i, 1,3)
        ),
      substring(i, 4),
      sep = "",
      collapse = " ")
    gene_names <- c(gene_names,p)
  }

  # Create data frame for matching names below
  df2 <- data.frame(genes,gene_names) %>%
    mutate(genes = as.character(genes)) %>%
    dplyr::rename(ref_name = genes)

  # Left join to match with ref_name column
  # and strip ref column for file suffix
  df <- df %>%
    left_join(df2, by = "ref_name") %>%
    mutate(gene_names = as.character(gene_names),
           ref = gsub("(.*?)_amr_report.tsv", "\\1", ref))

@@

# Function that handles megares data and returns a data 
# frame with information on whether a gene is mutated 
# or not
create_mut_table <- function(df) {
  df <- df %>%
    mutate(ref_ctg_change = ifelse(ref_ctg_change == ".", "", ref_ctg_change),
           ref_nt = ifelse(ref_nt == ".", "", ref_nt),
           ctg_nt = ifelse(ctg_nt == ".", "", ctg_nt),
           ref_ctg_nt_change = paste(ref_nt, ctg_nt, sep = "-"),
           mutation = paste(ref_ctg_change, ref_ctg_nt_change, sep = "/"),
           mutation = ifelse(mutation == "/-", "0", mutation)) %>%
    select(ref, gene_names, flag, mutation) %>%
    filter(flag %in% flag_selection) %>%
    mutate(id = 1:n()) %>%
    spread(gene_names, mutation) %>%
    group_by(ref) %>%
    summarise_all(funs(func_paste)) %>%
    select(-c(id, flag)) %>%
    gather(gene, mut, -ref) %>%
    mutate(mut = ifelse(mut == "" | mut == "." | is.na(mut) == TRUE, 0, mut),
           result = ifelse(mut != 0, 1, 0),
           result = as.integer(result),
           type = "mut",
           nt_change = gsub("[^,]+?/([^,]+?)", "\\1", mut),
           mut = gsub("([^/]+)/[^,]+", "\\1", mut))
  return(df)
}

# Corrects for mutations in QRDR for gyrA, gyrB, parC and parE genes.
# The function returns columns of 1/0 values for whether or not the 
# mutations reported by ARIBA is within the QRDR in the gene:
# gyrA: AA 67 - 106
# gyrB: AA 333 - 481
# ParC: AA 51 - 170
# parE: AA 366 - 523
# https://www.ncbi.nlm.nih.gov/pmc/articles/PMC1142146/pdf/cjvr68pg229.pdf
fix_gyr_par_results <- function(df) {
  df <- df %>%
    # control for mutation within QRDR for gyrA
    mutate(gyrA_result = mut %>% 
             str_extract_all("\\d+") %>% # from reprex package
             map(as.integer) %>% # converts all to integer
             # returns TRUE/FALSE whether value is within range or not
             map_lgl(~ any(.x >= 67L & .x <= 106L)), 
           gyrA_result = if_else(gene != "gyrA", NA, gyrA_result),
           gyrA_result = as.integer(gyrA_result),
           # control for mutation within QRDR for gyrB
           gyrB_result = mut %>%
             str_extract_all("\\d+") %>%
             map(as.integer) %>%
             map_lgl(~ any(.x >= 333L & .x <= 481L)),
           gyrB_result = if_else(gene != "gyrB", NA, gyrB_result),
           gyrB_result = as.integer(gyrB_result),
           # control for mutation within QRDR for parC
           parC_result = mut %>% 
             str_extract_all("\\d+") %>% 
             map(as.integer) %>% 
             map_lgl(~ any(.x >= 51L & .x <= 170L)),
           parC_result = if_else(gene != "parC", NA, parC_result),
           parC_result = as.integer(parC_result),
           # control for mutation within QRDR for parE
           parE_result = mut %>% 
             str_extract_all("\\d+") %>% 
             map(as.integer) %>% 
             map_lgl(~ any(.x >= 366L & .x <= 523L)),
           parE_result = if_else(gene != "parE", NA, parE_result),
           parE_result = as.integer(parE_result)) %>%
    # mergeresults from all four genes into one column
    mutate(result_gyr_par = case_when(gene == "gyrA" ~ gyrA_result,
                                      gene == "gyrB" ~ gyrB_result,
                                      gene == "parC" ~ parC_result,
                                      gene == "parE" ~ parE_result)) %>%
    # merge results from other genes with the four genes
    mutate(result_total = ifelse(gene %in% c("gyrA","gyrB","parC","parE"),
                                 result_gyr_par, result)) %>%
    select(-c(gyrA_result,
              gyrB_result,
              parC_result,
              parE_result,
              result_gyr_par,
              result))
  return(df)
}

# Function that handles resfinder data and returns a data frame with 
# presence/absence for acquired genes                           
create_acquired_table <- function(df) {
  df <- df %>%
    select(ref, gene_names, flag, ref_ctg_change) %>%
    mutate(id = 1:n()) %>%
    # Filter flags
    filter(flag %in% flag_selection) %>%
    spread(gene_names, ref_ctg_change) %>%
    group_by(ref) %>%
    summarise_all(funs(func_paste)) %>%
    mutate(ref = gsub("^(.*?)_amr_report.tsv$", "\\1", ref)) %>%
    select(-c(id, flag)) %>%
    gather(gene, result,-ref) %>%
    mutate(result_total = ifelse(result == "", 0, 1),
           result_total = as.character(result_total),
           type = "gene") %>%
    select(-result)
  return(df)
}

# Function that returns a filtered dataframe based 
# on the vector "mut_genes"
filter_mut_table <- function(df) {
  # Identify unique genes in column
  report_genes <- unique(df$gene)
  grep_genes <- c()

  # match names from name vector to column entries
  for (gene in mut_genes) {
    for (g in report_genes) {
      if (grepl(gene, g, ignore.case = T) == TRUE) {
        grep_genes <- c(grep_genes, g)
      }
    }
  }

  # Filter data frame based on values in grep_genes
  df <- df %>%
    filter(gene %in% grep_genes) %>%
    mutate(gene = as.character(gene))
  return(df)
}

# Function that returns a filtered dataframe based
# on the string "ac_genes"
filter_acquired_table <- function(df) {
  # Identify unique genes in column
  report_genes <- unique(df$gene)
  grep_genes <- c()

  # match names from vector to column entries
  for (gene in ac_genes) {
    for (g in report_genes) {
      if (grepl(gene, g, ignore.case = T) == TRUE) {
        grep_genes <- c(grep_genes, g)
      }
    }
  }

  # Filter data frame based on values in grep_genes
  df <- df %>%
    filter(gene %in% grep_genes) %>%
    mutate(gene = gsub("_", "", gene),
           gene = as.character(gene))
  return(df)
}

# calculates percentage of present/absent mutations and genes
calc_stats <- function(df) {
  df <- df %>%
    # Count Presence/Absence of each gene
    group_by(gene, result_total) %>%
    dplyr::count() %>%
    ungroup() %>%
    # Convert binary to "Present" and "Absent"
    mutate(result_total = if_else(result_total == 1, 
                                  "Present", "Absent")) %>%
    # Spread to separate Present and Absent columns
    spread(result_total, n, fill = 0) %>%
    rowwise() %>%
    # Calculate percentages
    mutate(Total = Present + Absent,
           Percent = round(Present/Total*100, 1))
  return(df)
}

# calculates how many mutations are present in the genes
# gyrA, gyrB, parC and parE
calc_no_of_mut <- function(df) {
  df1 <- df %>%
    # Filter out all other genes
    filter(gene %in% c("gyrA","parC","gyrB","parE")) %>%
    select(-nt_change) %>%
    # Count how many mutations in each gene per row
    mutate(entries = sapply(strsplit(.$mut, ","),
                            FUN = function(x) {length(x)})) %>%
    # Separate each mutation to own column
    separate(mut, into = as.character(c(1:max(.$entries)))) %>%
    select(-entries) %>%
    # Gather all mutations together into one column,
    # to get one row per mutation
    gather(id, mut, -c(ref, gene, result, type))

  # Correct for QRDR mutations in the four genes
  df2 <- fix_gyr_par_results(df1)

  # Generate final table
  df3 <- df2 %>%
    # Filter out mutations outide QRDR
    mutate(test = ifelse(mut != "0" & result_total == 0, 0, 1)) %>%
    filter(test == 1) %>%
    # Create own column for each gene
    spread(gene, mut) %>%
    group_by(ref) %>%
    # Summarise all columns based on unique values
    summarise_all(funs(func_paste)) %>%
    # Remove "0" values for correct counts below
    mutate_at(.vars = vars(c("gyrA","gyrB","parC","parE")),
              .funs = function(x) ifelse(x == "0", "", x)) %>%
    # Count mutations in each column entry
    mutate(mut_gyrA = sapply(strsplit(.$gyrA, ","),
                             FUN = function(x) {length(x)}),
           mut_gyrB = sapply(strsplit(.$gyrB, ","),
                             FUN = function(x) {length(x)}),
           mut_parC = sapply(strsplit(.$parC, ","),
                             FUN = function(x) {length(x)}),
           mut_parE = sapply(strsplit(.$parE, ","),
                             FUN = function(x) {length(x)})) %>%
    select(-c(type, id, result_total, test)) %>%
    # Reinsert "0" values 
    mutate_at(.vars = vars(c("gyrA","gyrB","parC","parE")),
              .funs = function(x) ifelse(x == "", "0", x))
  return(df3)
}

@@

# -------------------------- Analysis ---------------------------
# Import data
mut_data <- get_ariba_data(megares_report_loc)
acquired_data <- get_ariba_data(resfinder_report_loc)

@@

# Check flags
mut_flags <- check_flags(clean_mut_data) %>%
  left_join(id_data, by = "ref") %>%
  select(id, everything(), -ref) %>%
  # Filters out unwanted samples
  filter(id %not_in% ex_samples)

acquired_flags <- check_flags(clean_acquired_data) %>%
  left_join(id_data, by = "ref") %>%
  select(id, everything(), -ref) %>%
  filter(id %not_in% ex_samples)

mut_table <- create_mut_table(clean_mut_data)

mut_table_fixed <- fix_gyr_par_results(mut_table) %>%
  left_join(id_data, by = "ref") %>%
  select(id, everything(), -ref) %>%
  filter(id %not_in% ex_samples)

mut_filtered <- filter_mut_table(mut_table_fixed)

acquired_table <- create_acquired_table(clean_acquired_data) %>%
  left_join(id_data, by = "ref") %>%
  select(id, everything(), -ref) %>%
  filter(id %not_in% ex_samples)

@@

mut_stats <- calc_stats(mut_filtered)
mut_report <- create_mut_report(mut_filtered)

acquired_stats <- calc_stats(acquired_filtered)
acquired_report <- create_acquired_report(acquired_filtered)

mut_quant <- suppressWarnings(calc_no_of_mut(mut_table)) %>%
  # Correct for erroneous entry from ARIBA
  mutate(parC = ifelse(parC == "S58I", "S80I", parC)) %>%
  left_join(id_data, by = "ref") %>%
  select(id, everything(), -ref) %>%
  filter(id %not_in% ex_samples)

@@

# Tables
## Isolate data
This table presents the isolates and their accompanying metadata.
```{r}
isolate_data_complete <- isolate_data %>%
  # combine with read data
  left_join(no_of_reads, by = "id")

@@

## Isolates per species
This table presents how many isolates were included per species.
```{r}
per_species <- isolate_data %>%
  group_by(species) %>%
  dplyr::count()
